---
title: "Replication of 'Why Do We Hate Hypocrites? Evidence for a Theory of False Signaling' by Jillian J. Jordan et al. (2017, Psychological Science)"
author: "Andrew Joo Hun Nam (ajhnam@stanford.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
---

<!-- Replication reports should all use this template to standardize reporting across projects.  These reports will be public supplementary materials that accompany the summary report(s) of the aggregate results. -->

##Introduction

'Why Do We Hate Hypocrites? Evidence for a Theory of False Signaling' by Jordan et al. explored the possible reasons why hypocrisy is treated with notably high disapproval. The authors published the research along with the data and materials made publically accessible on Open Science Framework which makes the study a strong candidate for replicability. It is hoped that a study following more rigorous scientific protocol will have a higher success in replication.

Through the 5 studies, Jordan et al. suggest that condemnation acts as a strong signal for virtuous character, inviting greater disdain upon transgression than verbal statements about own moral character. Their results also supported the theory that the response to hypocrisy is a response to feeling manipulated as demonstrated in their comparison to hypocrisy to 'honest' hypocrisy (transgression and condemnation but also with confession) where the disapproval disappeared. In this replication study, I replicate the third study by Jordan et al.

<!-- In this replication study, I replicate 3 out of the 5 studies by Jordan et al., specifically studies 1, 3, and 4, which are re-indexed in this study as studies 1, 2, and 3.  -->


##Methods

###Power Analysis

A sample size of 451 participants yielded the original study an effect size of f = 0.3443 (d = 0.6887). To achieve 80% power for detecting the effect size at $\alpha = .05$, the study would require a sample size of 85 participants. 110 participants would yield a 90% power and 134 participants for a 95% power.

###Planned Sample

In this replication study, I sampled across 85 participants to achieve an 80% power to detect the effect. The study only analyzed subjects who had unique IP addresses and evaluated all target characters.

###Materials

Vignettes and response scales were followed exactly in accordance with the original study.

####Vignettes

Subjects evaluted the following four scenarios in a random order:

#####Music Scenario (Person of Interest: Becky)

Becky and her friend Amanda are discussing a mutual acquaintance. Amanda mentions that the acquaintance often downloads music illegally from the Internet.

Hypocrisy condition: Becky says that she thinks it is morally wrong to download music illegally from the Internet.

Liar condition: Becky says that she doesn't download music illegally from the Internet.

Control: No sentence presented in the control transgressor condition.

Shortly after their conversation, Becky goes online, and downloads music illegally.

#####Jury Scenario (Person of Interest: Jennifer)

Jennifer and her friend Rose are discussing a mutual acquaintance. Rose mentions that the acquaintance recently tried to get out of jury duty. 

Hypocrisy condition: Jennifer says that she thinks it is morally wrong to try to get out of jury duty.

Liar condition: Jennifer says that she doesn't try to get out of jury duty.

Control: No sentence presented in the control transgressor condition

Shortly after their conversation, Jennifer gets called for jury duty, and tries to get out of it.

#####Phone Call Scenario (Person of Interest: Bruce)

Bruce and his friend are Zach are discussing a mutual acquaintance. Zach mentions that the acquaintance often ignores his mother's phone calls. 

Hypocrisy condition: Bruce says that he thinks it is morally wrong to ignore your mother's phone calls.

Liar condition: Bruce says that he doesn't ignore his mother's phone calls.

Control: No sentence presented in the control transgressor condition.

Shortly after their conversation, Bruce notices that his mother is calling, and ignores the call.

#####Printing Scenario (Person of Interest: Kevin)

Kevin and his friend Jack are discussing a mutual acquaintance. Jack mentions that the acquaintance often uses a lot of paper by printing documents single-sided.

Hypocrisy condition: Kevin says that he thinks it is morally wrong to use a lot of paper by printing documents single-sided.

Liar condition: Kevin says that he doesn’t use a lot of paper by printing documents single-sided.

No sentence presented in the control transgressor condition.

Shortly after their conversation, Kevin has a large document to print, and uses a lot of paper by printing it single-sided.

####Response Questions

The participants were presented a sliding scale with no numerical values.

1. How trustworthy do you think [Person of Interest] is? - Scale from "Not at all trustworthy" to "Very trustworthy"

2. How hypocritical do you think [Person of Interest] is? - Scale from "Not at all hypocritical" to "Very hypocritical"

3. How much do you like [Person of Interest]? - Scale from "Not at all" to "Very much"

4. How honest do you think [Person of Interest] is? - Scale from "Not at all honest" to "Very honest"

5. How good a person do you think Becky is? - Scale from "Not at all good" to "Very good"

###Procedure	

The procedure was followed exactly in accordance with the original study.

> We created vignettes in which a target character discusses an acquaintance’s moral transgression and then privately goes on to engage in the same transgression. We manipulated whether, in addition to committing the transgression, the target was a hypocrite, a (direct) liar, or neither.

> We again presented subjects with vignettes and asked them to evaluate the target characters in the vignettes. In a three-condition, between-subjects design, we manipulated whether, before engaging in the relevant moral violation, the target character (a) condemned the violation (hypocrisy condition), (b) directly stated that he or she did not engage in the violation (liar condition), or (c) said nothing (control-transgressor condition). We predicted that subjects would evaluate hypocrites as both worse than control transgressors and worse than liars.

> Each vignette described a conversation between two characters: a target (whom subjects would later evaluate) and a friend (whom subjects would not evaluate). In all conditions, this conversation began with the target and the friend discussing a mutual acquaintance. In this discussion, the friend mentioned that the mutual acquaintance often engaged in a particular moral transgression.

> In the hypocrisy condition, the target responded to the friend by condemning the transgression. In contrast, in the liar condition, the target responded by directly stating that he or she did not engage in the relevant transgression. Finally, in the control-transgressor condition, we did not include any information about a response from the target. Shortly after this conversation ended, in all conditions, the target went on to commit the relevant violation.

> For example, here is the full text for a scenario about downloading music illegally. In this scenario, Becky is the target character, and Amanda is the friend. In all conditions, the vignette began as follows:

> Becky and her friend Amanda are discussing a mutual acquaintance. Amanda mentions that the acquaintance often downloads music illegally from the Internet.

> In the hypocrisy condition, the scenario continued,

> Becky says that she thinks it is morally wrong to download music illegally from the Internet. Shortly after their conversation, Becky goes online, and downloads music illegally.

> In the liar condition, the scenario instead continued,

> Becky says that she doesn’t download music illegally from the Internet. Shortly after their conversation, Becky goes online, and downloads music illegally.

> Finally, in the control-transgressor condition, nothing was said about Becky’s opinion or behavior, and the scenario simply ended with

> Shortly after their conversation, Becky goes online, and downloads music illegally.

> Each subject was presented with four vignettes (in random order), about downloading music illegally, evading jury duty, ignoring phone calls from one’s mother, and wasting paper by printing documents single-sided.

> After reading each vignette, subjects evaluated the target. We asked subjects to rate how good a person the target was, how much they liked the target, how honest they thought the target was, and how trustworthy they thought the target was. Further, as a manipulation check, and to conduct an exploratory investigation into how subjects conceptualize “hypocrisy,” we asked them to rate how hypocritical the target was. These five dependent measures were presented in random order for each vignette. Subjects responded to each item on a sliding scale, with anchors reading not at all [trait] to very [trait] (e.g., not at all trustworthy to very trustworthy). The sliding scales did not have any numerical labels, but responses were translated to scores ranging from 0 (not at all) to 100 (very).

> We found high interitem reliability among our four primary individual dependent measures (i.e., excluding our hypocrisy variable, which was used as a manipulation check; α = .96). Thus, as in Studies 1 and 2, we averaged these ratings to create a composite scale representing positive evaluations of the target and used this composite as our dependent variable in the analyses reported here (analyses investigating the individual dependent measures are reported in the Supplemental Material). We note that including the hypocrisy measure in our composite variable did not qualitatively affect our conclusions; however, results from the hypocrisy measure followed a somewhat different pattern than results from the other measures, and provide insight into the ways that subjects use the term “hypocrisy” (see the Supplemental Material).

###Analysis Plan

The analysis was followed exactly in according to the original study.

> To test our prediction, we conducted a one-way ANOVA
investigating the effect of condition on positive evaluations
of the targets across the vignettes.

More specifically, the scale ratings were averaged for each participant across each question and vignette as a composite measure for evaluation of targets. Then a one-way ANOVA was performed regressing the condition (hypocrite, liar, or control transgressor) to the evaluation of targets measure.

###Differences from Original Study

The original authors also collected demographics data regarding age, gender, education, income, generaltrust, reason for their choices, and previous experience in similar studies. These were excluded in the replication. These variables, recorded in the supplementary materials and open-sourced dataset, record minimum details for implementation (e.g. it's unclear whether these questions were presented before or after the vignette-response sections). Moreover, the authors report no significant interactions between the conditions and demographics data.

### Methods Addendum (Post Data Collection)

You can comment this section out prior to final report with data collection.

#### Actual Sample
Sample size, demographics, data exclusions based on rules spelled out in analysis plan

#### Differences from pre-data collection methods plan
Any differences from what was described as the original plan, or “none”.


##Results

### Data preparation

Data preparation following the analysis plan.

```{r include=F}
####Load Relevant Libraries and Functions
library("readxl")
library(dplyr)
library(tidyr)
library(Rmisc)
library(broom)
####Import data

# Using original data
orig_data <- read_excel("../original_paper/original_data.xls", sheet=3)

head(orig_data)
```

### Direct Regression From Original Data
```{r}
summary(lm(overall ~ condition, data=orig_data))
```

### Full Reproduction From Original Data
```{r}
```

### Using Pilot Data
```{r}
####Import data
filename <- "../pilot_a/data/Psych 251 Pilot A 2_November 2, 2018_11.23.csv"
data <- read.csv(filename)

###Data Preparation
data <- data[-c(1,2), ]
data <- data %>%
  mutate(subject = 1:nrow(data)) %>%
  select(
    subject, starts_with("Q"), condition
  ) %>%
  drop_na() %>%
  mutate_at(vars(starts_with("Q")),funs(as.numeric))


colnames(data)[colnames(data)=="Q3_1"] <- "trustmu_1"
colnames(data)[colnames(data)=="Q5_1"] <- "hypmu_1"
colnames(data)[colnames(data)=="Q7_1"] <- "likemu_1"
colnames(data)[colnames(data)=="Q9_1"] <- "goodmu_1"
colnames(data)[colnames(data)=="Q11_1"] <- "honest_mu1"
colnames(data)[colnames(data)=="Q13_1"] <- "trustjd_1"
colnames(data)[colnames(data)=="Q19_1"] <- "hypjd_1"
colnames(data)[colnames(data)=="Q27_1"] <- "likejd_1"
colnames(data)[colnames(data)=="Q31_1"] <- "goodjd_1"
colnames(data)[colnames(data)=="Q37_1"] <- "honestjd_1"
colnames(data)[colnames(data)=="Q15_1"] <- "trustpc"
colnames(data)[colnames(data)=="Q21_1"] <- "hyppc"
colnames(data)[colnames(data)=="Q25_1"] <- "likepc"
colnames(data)[colnames(data)=="Q33_1"] <- "goodpc"
colnames(data)[colnames(data)=="Q39_1"] <- "honestpc"
colnames(data)[colnames(data)=="Q41_1"] <- "honestpd_1"
colnames(data)[colnames(data)=="Q17_1"] <- "trustpd_1"
colnames(data)[colnames(data)=="Q23_1"] <- "hyppd_1"
colnames(data)[colnames(data)=="Q35_1"] <- "goodpd_1"
colnames(data)[colnames(data)=="Q29_1"] <- "likepd_1"

head(data)

#### Prepare data for analysis - create columns etc.
data <- data %>%
  mutate(
    like = rowMeans(select(., starts_with("like"))),
    good = rowMeans(select(., starts_with("good"))),
    trust = rowMeans(select(., starts_with("trust"))),
    hyp = rowMeans(select(., starts_with("hyp"))),
    honest = rowMeans(select(., starts_with("honest"))),
  )
data <- data %>%
  mutate(overall = rowMeans(select(., like, good, trust, hyp, honest)))
```

```{r}
labels <- c("Control","Hypocrite","Liar")
barplot(c(4.625, 4.625-.075, 4.625-.025), names.arg=labels, xlab="Condition",ylab="Rating")

```

### Confirmatory analysis

The analyses as specified in the analysis plan.
```{r}
summary(lm(overall ~ condition, data=orig_data))
```


*Side-by-side graph with original graph is ideal here*

###Exploratory analyses

Any follow-up analyses desired (not required).  

## Discussion

### Summary of Replication Attempt

Open the discussion section with a paragraph summarizing the primary result from the confirmatory analysis and the assessment of whether it replicated, partially replicated, or failed to replicate the original result.  

### Commentary

Add open-ended commentary (if any) reflecting (a) insights from follow-up exploratory analysis, (b) assessment of the meaning of the replication (or not) - e.g., for a failure to replicate, are the differences between original and present study ones that definitely, plausibly, or are unlikely to have been moderators of the result, and (c) discussion of any objections or challenges raised by the current and original authors about the replication attempt.  None of these need to be long.
